Summary of antibiotic consumption

### 2017-02-22 hicks-summary/new-hicks/hicks-summary2.Rmd
new hicks

- Change `consumption_by_class` to `rx_by_class`

```{r setup, echo=FALSE}
# Group values in x into ranges determined by upper_limits.
# Values below the lowest limit are named as the lowest limit, and so forth.
# You should probably put Inf as the last limit.
bounds_group = function(x, upper_limits) {
  range_strings = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('(%i) < %.1f', i, upper_limits[i])
    } else {
      sprintf('(%i) %.1f -- %.1f', i, upper_limits[i - 1], upper_limits[i])
    }
  })
  formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
                        range_strings, upper_limits)
  do.call(case_when, formula_list)
}

summarize_by = function(x, bene, by) {
  group_by_(x, by) %>%
    summarize(n_claims=sum(n_claims)) %>%
    left_join(bene %>% group_by_(by) %>% summarize(n_ppl=n()), by=by) %>%
    mutate(cpkp=n_claims*1000/n_ppl) %>%
    select_(by, 'cpkp')
}
```

```{r load_db, echo=FALSE}
library(knitr)

db = src_sqlite('../../abx.sqlite')

regions = tbl(db, 'census_regions') %>%
  select(state, region)

abx_class = tbl(db, 'abx_class') %>%
  select(abx, abx_class)
```

```{r analyze_data, cache=TRUE, echo=FALSE}
analyze = function(this_year) {
  bene = tbl(db, sprintf('bene_%i', this_year)) %>%
    filter(!is.na(state)) %>%
    left_join(regions, by='state') %>%
    mutate(age=as.integer(age)) %>%
    compute

  pde = tbl(db, sprintf('abx_pde_%i', this_year)) %>%
    left_join(abx_class, by='abx') %>%
    left_join(bene, by='bene') %>%
    mutate(n_claims=as.integer(n_claims)) %>%
    compute

  total_n_bene = bene %>% count() %>% head(1) %>% collect %$% n

  # claims per 1,000 people by abx class
  consumption_by_class = pde %>%
    group_by(abx_class) %>%
    summarize(n_claims=sum(n_claims)) %>%
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    mutate(year=this_year) %>%
    select(abx_class, cpkp, year) %>% 
    collect

  # the top few inidividual antibiotics
  top_abx = pde %>%
    group_by(abx) %>%
    summarize(n_claims=sum(n_claims)) %>% 
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    mutate(year=this_year) %>%
    select(abx, cpkp, year) %>%
    collect

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(pde, bene, 'sex') %>% mutate(year=this_year) %>% collect

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  age_group_f = function(x) bounds_group(x, c(65, 75, 85, Inf))
  bene_age = bene %>%
    filter(!is.na(age)) %>%
    collect(n=Inf) %>% 
    mutate(age_group=age_group_f(age)) %>%
    count(age_group) %>%
    rename(n_bene=n)

  claims_by_age = pde %>%
    collect(n=Inf) %>%
    mutate(age_group=age_group_f(age)) %>%
    group_by(age_group) %>%
    summarize(n_claims=sum(n_claims)) %>%
    left_join(bene_age, by='age_group') %>%
    mutate(cpkp=n_claims*1000/n_bene, year=this_year) %>%
    select(age_group, cpkp, year)

  # then by census region
  claims_by_region = summarize_by(pde, bene, 'region') %>%
    mutate(year=this_year) %>% collect

  # claims by state (broken into sextiles)
  claims_by_state = summarize_by(pde, bene, 'state') %>%
    collect(n=Inf) %>%
    mutate(sextile=ntile(cpkp, 6), year=this_year)

  list(consumption_by_class=consumption_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       claims_by_region=claims_by_region,
       claims_by_state=claims_by_state)
}

@@

claims_by_class = data_by_field('consumption_by_class')

top_classes = claims_by_class %>%
  filter(year==2011) %>%
  arrange(desc(cpkp)) %>%
  head(9) %$%
  abx_class

claims_by_class %<>%
  mutate(abx_class=if_else(abx_class %in% top_classes, abx_class, 'other')) %>%
  group_by(year, abx_class) %>%
  summarize(cpkp=sum(cpkp)) %>%
  spread(year, cpkp) %>%
  # add this is_other variable to make sure that 'other' appears in the bottom row
  # swo> could also do this with an ordered factor
  mutate(is_other=abx_class=='other') %>%
  arrange(is_other, desc(`2011`)) %>%
  select(-is_other) %>%
  kable

@@

claims_by_region = data_by_field('claims_by_region') %>%
  filter(!is.na(region)) %>%
  spread(year, cpkp) %>%
  kable

>>>>

```{r setup, echo=FALSE}
library(assertive)

# Group values in x into ranges determined by upper_limits.
# Values below the lowest limit are named as the lowest limit, and so forth.
# You should probably put Inf as the last limit.
bounds_group = function(x, upper_limits) {
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('< %.1f', upper_limits[i])
    } else {
      sprintf('>= %.1f and < %.1f', upper_limits[i - 1], upper_limits[i])
    }
  })
  formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
                        levels, upper_limits)
  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

summarize_by = function(usage, bene, by) {
  denom = group_by_(bene, by) %>%
    summarize(n_ppl=n())

  group_by_(usage, by) %>%
    summarize(n_claims=n()) %>%
    left_join(denom, by=by) %>%
    mutate(cpkp=n_claims*1000/n_ppl) %>%
    select_(by, 'cpkp')
}
```

```{r load_db, echo=FALSE, message=FALSE}

regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
  select(state, region)

antibiotic_class = read_tsv('../../abx_class.tsv') %>%
  select(antibiotic, antibiotic_class) %>%
  distinct

assert_has_no_duplicates(antibiotic_class$antibiotic)
```

```{r analyze_data, cache=TRUE, echo=FALSE, message=FALSE}
analyze = function(this_year) {
  bene = read_tsv(sprintf('../../bene_%i.tsv', this_year)) %>%
    mutate_at(vars(age, plan_coverage_months), as.integer) %>%
    filter(!is.na(state), plan_coverage_months == 12, age >= 65) %>%
    left_join(regions, by='state')

  assert_has_no_duplicates(bene$bene_id)
  total_n_bene = nrow(bene)

  pde = read_tsv(sprintf('../../pde_%i.tsv', this_year)) %>%
    mutate_at(vars(days_supply), as.integer) %>%
    left_join(antibiotic_class, by='antibiotic')

  usage = pde %>%
    inner_join(bene, by='bene_id') # inner join to keep only matching rows

  # usage only has PDEs from beneficiaries in our list,
  # so there should be fewer
  is_greater_than_or_equal_to(nrow(pde), nrow(usage))

  # claims per 1,000 people by abx class
  claims_by_class = usage %>%
    group_by(antibiotic_class) %>%
    summarize(n_claims=n()) %>%
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    select(antibiotic_class, cpkp) %>%
    mutate(year=this_year)

  # the top few inidividual antibiotics
  top_abx = usage %>%
    group_by(antibiotic) %>%
    summarize(n_claims=n()) %>%
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    select(antibiotic, cpkp) %>% 
    mutate(year=this_year)

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(usage, bene, 'sex') %>%
    mutate(year=this_year)

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  age_group_f = function(x) bounds_group(x, c(65, 75, 85, Inf))
  claims_by_age = summarize_by(
    usage %>% mutate(age_group=age_group_f(age)),
    bene %>% mutate(age_group=age_group_f(age)),
    'age_group') %>%
    mutate(year=this_year)

  # then by census region
  claims_by_region = summarize_by(usage, bene, 'region') %>%
    mutate(year=this_year)

  # claims by state (broken into sextiles)
  claims_by_state = summarize_by(usage, bene, 'state') %>%
    mutate(sextile=ntile(cpkp, 6), year=this_year)

  list(claims_by_class=claims_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       claims_by_region=claims_by_region,
       claims_by_state=claims_by_state)
}

@@

claims_by_class = data_by_field('claims_by_class')

top_classes = claims_by_class %>%
  filter(year==2011) %>%
  arrange(desc(cpkp)) %>%
  head(9) %$%
  antibiotic_class %>%
  c('other')

claims_by_class %<>%
  mutate(antibiotic_class=if_else(antibiotic_class %in% top_classes,
                                  antibiotic_class,
                                  'other')) %>%
  group_by(year, antibiotic_class) %>%
  summarize(cpkp=sum(cpkp)) %>%
  spread(year, cpkp) %>%
  mutate(antibiotic_class=factor(antibiotic_class, levels=top_classes, ordered=TRUE)) %>%
  arrange(antibiotic_class) %>%
  kable

@@

claims_by_region = data_by_field('claims_by_region') %>%
  filter(!is.na(region)) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>%
  kable
